\documentclass{article}
\usepackage{parskip}
\usepackage[margin=1in]{geometry}
\usepackage{mathptmx}
\usepackage{courier}
\usepackage[scaled=.92]{helvet}
\usepackage[T1]{fontenc}

\pagestyle{empty}
\begin{document}
\today

Raghav Kansal has my permission to include the following papers, of which
I was a co-author, in his doctoral dissertation.

\begin{enumerate}
    \item A. Li*, V. Krishnamohan*, R. Kansal, J. Duarte, R. Sen, S. Tsan, and Z. Zhang, “Induced generative adversarial particle transformers”, NeurIPS ML4PS Workshop (2023), arXiv:2312.04757.
    \item R. Kansal, C. Pareja, Z. Hao, and J. Duarte, “JetNet: A Python package for accessing open datasets and benchmarking machine learning methods in high energy physics”, JOSS 8, 5789 (2023).
    \item Z. Hao, R. Kansal, J. Duarte, and N. Chernyavskaya, “Lorentz group equivariant autoencoders”, Eur. Phys. J. C 83, 485 (2023), arXiv:2212.07347.
    \item R. Kansal, A. Li, J. Duarte, N. Chernyavskaya, M. Pierini, B. Orzari, and T. Tomei, “Evaluating generative models in high energy physics”, Phys. Rev. D 107, 076017 (2023), arXiv:2211.10295.
    \item R. Kansal, J. Duarte, H. Su, B. Orzari, T. Tomei, M. Pierini, M. Touranakou, J.-R. Vlimant, and D. Gunopulos, “Particle Cloud Generation with Message Passing Generative Adversarial Networks”, NeurIPS (2021), arXiv:2106.11535.
    \item S. Tsan, R. Kansal, A. Aportela, D. Diaz, J. Duarte, S. Krishna, F. Mokhtar, J.-R. Vlimant, and M. Pierini, “Particle graph autoencoders and differentiable, learned energy mover's distance”, NeurIPS ML4PS Workshop (2021), arXiv:2111.12849.
    \item R. Kansal, J. Duarte, B. Orzari, T. Tomei, M. Pierini, M. Touranakou, J.-R. Vlimant, and D. Gunopulos, “Graph generative adversarial networks for sparse data generation in high energy physics”, NeurIPS ML4PS Workshop (2020), arXiv:2012.00173.
\end{enumerate}

\vskip3\baselineskip
----------------------------------------\\
Prof. Javier Duarte
\end{document}
