\noindent
\begin{cv}{}
\begin{cvlist}{}
\item[2019] Bachelor of Science in Physics and Computer Engineering\\
            University of California San Diego\\
            \textit{Summa cum laude}
\item[2019] CERN Openlab Summer Student
\item[2019] IRIS-HEP Fellow
\item[2019-2024] Graduate Student Researcher\\University of California San Diego
\item[2021-2022] Artificial Intelligence Fellow\\Fermilab LHC Physics Center
\item[2023-2024] Graduate Scholar\\Fermilab LHC Physics Center
\item[2024] Doctor of Philosophy in Physics\\University of California San Diego
\item[2024-] AI/Schmidt Postdoctoral Scholar Research Associate\\California Institute of Technology and Fermilab
\end{cvlist}
\end{cv}

% \clearpage
\publications

\noindent
\textit{Note: as a member of the CMS collaboration, I have been an author on all CMS papers since 2019.
The following includes only the CMS publications to which I made significant contributions during my PhD.}
\\[0.1\baselineskip]
\begin{enumerate}[itemsep=\baselineskip]
    \item CMS Collaboration, “Search for Nonresonant Pair Production of Highly Energetic Higgs Bosons Decaying to Bottom Quarks and Vector Bosons”, in prep, CMS-HIG-23-012 (2023).
    \item CMS Collaboration, “Search for a massive scalar resonance decaying to a light scalar and a Higgs boson in the two b quarks and four light quarks final state”, in prep, CMS-B2G-23-007 (2023).
    \item A. Li*, V. Krishnamohan*, \textcolor{authorhighlight}{R. Kansal}, J. Duarte, R. Sen, S. Tsan, and Z. Zhang, “Induced generative adversarial particle transformers”, NeurIPS ML4PS Workshop (2023), \href{https://arxiv.org/abs/2312.04757}{arXiv:2312.04757}.
    \item \textcolor{authorhighlight}{R. Kansal}, C. Pareja, Z. Hao, and J. Duarte, “JetNet: A Python package for accessing open datasets and benchmarking machine learning methods in high energy physics”, \href{https://doi.org/10.21105/joss.05789}{\textbf{JOSS} 8, 5789} (2023).
    \item Z. Hao, \textcolor{authorhighlight}{R. Kansal}, J. Duarte, and N. Chernyavskaya, “Lorentz group equivariant autoencoders”, \href{https://doi.org/10.1140/epjc/s10052-023-11633-5}{\textbf{Eur. Phys. J. C} 83, 485} (2023), \href{https://arxiv.org/abs/2212.07347}{arXiv:2212.07347}.
    \item \textcolor{authorhighlight}{R. Kansal}, A. Li, J. Duarte, N. Chernyavskaya, M. Pierini, B. Orzari, and T. Tomei, “Evaluating generative models in high energy physics”, \href{https://doi.org/10.1103/PhysRevD.107.076017}{\textbf{Phys. Rev. D} 107, 076017} (2023), \href{https://arxiv.org/abs/2211.10295}{arXiv:2211.10295}.
    \item CMS Collaboration, “Search for Nonresonant Pair Production of Highly Energetic Higgs Bosons Decaying to Bottom Quarks”, \href{https://doi.org/10.1103/PhysRevLett.131.041803}{\textbf{Phys. Rev. Lett.} 131, 041803} (2023), \href{https://arxiv.org/abs/2205.06667}{arXiv:2205.06667}.
    \item \textcolor{authorhighlight}{R. Kansal}, J. Duarte, H. Su, B. Orzari, T. Tomei, M. Pierini, M. Touranakou, J.-R. Vlimant, and D. Gunopulos, “Particle Cloud Generation with Message Passing Generative Adversarial Networks”, NeurIPS (2021), \href{https://arxiv.org/abs/2106.11535}{arXiv:2106.11535}.
    \item F. Mokhtar, \textcolor{authorhighlight}{R. Kansal}, and J. Duarte, “Do graph neural networks learn traditional jet substructure?”, NeurIPS ML4PS Workshop (2022), \href{https://arxiv.org/abs/2211.09912}{arXiv:2211.09912}.
    \item M. Touranakou, N. Chernyavskaya, J. Duarte, D. Gunopulos, \textcolor{authorhighlight}{R. Kansal}, B. Orzari, M. Pierini, T. Tomei, and J.-R. Vlimant, “Particle-based fast jet simulation at the LHC with variational autoencoders”, \href{https://doi.org/10.1088/2632-2153/ac7c56}{\textbf{Machine Learning: Science and Technology} 3, 035003} (2022), \href{https://arxiv.org/abs/2203.00520}{arXiv:2203.00520}.
    \item A. Apresyan, D. Diaz, J. Duarte, S. Ganguly, \textcolor{authorhighlight}{R. Kansal}, N. Lu, C. M. Suarez, S. Mukherjee, C. Peña, B. Sheldon, and S. Xie, “Improving Di-Higgs Sensitivity at Future Colliders in Hadronic Final States with Machine Learning”, Contribution to Snowmass Summer Study (2022), \href{https://arxiv.org/abs/2203.07353}{arXiv:2203.07353}.
    \item Y. Chen et al., “A FAIR and AI-ready Higgs boson decay dataset”, \href{https://doi.org/10.1038/s41597-021-01109-0}{\textbf{Sci. Data} 9, 31} (2021), \href{https://arxiv.org/abs/2108.02214}{arXiv:2108.02214}.
    \item F. Mokhtar, \textcolor{authorhighlight}{R. Kansal}, D. Diaz, J. Duarte, J. Pata, M. Pierini, and J.-R. Vlimant, “Explaining machine-learned particle-flow reconstruction”, NeurIPS ML4PS Workshop (2021), \href{https://arxiv.org/abs/2111.12840}{arXiv:2111.12840}.
    \item S. Tsan, \textcolor{authorhighlight}{R. Kansal}, A. Aportela, D. Diaz, J. Duarte, S. Krishna, F. Mokhtar, J.-R. Vlimant, and M. Pierini, “Particle graph autoencoders and differentiable, learned energy mover’s distance”, NeurIPS ML4PS Workshop (2021), \href{https://arxiv.org/abs/2111.12849}{arXiv:2111.12849}.
    \item B. Orzari, T. Tomei, M. Pierini, M. Touranakou, J. Duarte, \textcolor{authorhighlight}{R. Kansal}, J.-R. Vlimant, and D. Gunopulos, “Sparse Data Generation for Particle-Based Simulation of Hadronic Jets in the LHC”, ICML LXAI Workshop (2021), \href{https://arxiv.org/abs/2109.15197}{arXiv:2109.15197}.
    \item \textcolor{authorhighlight}{R. Kansal}, J. Duarte, B. Orzari, T. Tomei, M. Pierini, M. Touranakou, J.-R. Vlimant, and D. Gunopulos, “Graph generative adversarial networks for sparse data generation in high energy physics”, NeurIPS ML4PS Workshop (2020), \href{https://arxiv.org/abs/2012.00173}{arXiv:2012.00173}.
\end{enumerate}
